---
title: "HW5 - Make Data Move"
author: ''
date: "05/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Explore the recent global developments with R
Today, you will load a filtered gapminder dataset -  with a subset of data on global development from 1952 - 2007 in increments of 5 years - to capture the period between the Second World War and the Global Financial Crisis. 

**Your task: Explore the data and visualise it in both static and animated ways, providing answers and solutions to 7 questions/tasks below.**

## Get the necessary packages
First, start with installing the relevant packages 'tidyverse', 'gganimate', and 'gapminder'.

```{r libraries, echo = FALSE}
library(tidyverse)
library(gganimate)
library(gapminder)
```

## Look at the data
First, see which specific years are actually represented in the dataset and what variables are being recorded for each country. Note that when you run the cell below, Rmarkdown will give you two results - one for each line - that you can flip between.
```{r}
unique(gapminder$year)
head(gapminder)
```
The dataset contains information on each country in the sampled year, its continent, life expectancy, population, and GDP per capita.

Let's plot all the countries in 1952.
```{r 1957}
theme_set(theme_bw())  # set theme to white background for better visibility

ggplot(subset(gapminder, year == 1952), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() 
```
We see an interesting spread with an outlier to the right. Answer the following questions, please:

Q1. Why does it make sense to have a log10 scale on x axis? 
# Q1: As you can see in the visualization above one dot is far away from the others. Therefore the graph would be very long if we did not have a log10 scale on the x axis. By having a log10 on the x axis the grapgh is easier to comprehend.

Q2. What country is the richest in 1952 (far right on x axis)? 

```{r, Q2}
gapminder %>% 
  filter(year == 1952) %>% # I want to see the numbers from 1952
  select(country, gdpPercap) %>% # I only want to look at the catagory that shows countries' prosperity
  arrange(desc(gdpPercap)) # I want to see the richest country first
```

# Q2: The richest country in 1952 was Kuwait.

You can generate a similar plot for 2007 and compare the differences
```{r 2007}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() 
```
The black bubbles are a bit hard to read, the comparison would be easier with a bit more visual differentiation.

Q3. Can you differentiate the continents by color and fix the axis labels?
```{r, Q3}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +  # I want to see the graph from 2007 and look at GDP and life expectancy
  geom_point(aes(color = continent)) + # I color the continents
  labs(x = "GDP") + # I title the x axis "GDP"" 
  labs(y = "lifeExp") # I title the y axis "lifeExp""

```


Q4. What are the five richest countries in the world in 2007?

```{r, Q4}
gapminder %>% 
  filter(year == 2007) %>% # I want to see the numbers from 2007
  select(country, gdpPercap) %>% # I only want to look at the catagory that shows countries' prosperity
  arrange(desc(gdpPercap)) %>% # I want to see the richest country first
  head(5) # I want to see the five richest countries first
```

# Q4: The five richest countries in the world in 2007 were Norway, Kuwait, Singapore, United States, and Ireland,



##  Make it move!

The comparison would be easier if we had the two graphs together, animated. We have a lovely tool in R to do this: the `gganimate` package. And there are two ways of animating the gapminder ggplot.

### Option 1: Animate using transition_states() 

The first step is to create the object-to-be-animated
```{r anim1 start}
anim <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10()  # convert x to log scale
anim
```

This plot collates all the points across time. The next step is to split it into years and animate it. This may take some time, depending on the processing power of your computer (and other things you are asking it to do). Beware that the animation might appear in the 'Viewer' pane, not in this rmd preview. You need to knit the document to get the viz inside an html file.

```{r anim1}
anim + transition_states(year, 
                      transition_length = 1,
                      state_length = 1)

```
Notice how the animation moves jerkily, 'jumping' from one year to the next 12 times in total. This is a bit clunky, which is why it's good we have another option. 


### Option 2 Animate using transition_time()

```{r packages}
install.packages("gifski")
install.packages("png")
library(gifski)
library(png)
```


This option smoothes the transition between different 'frames', because it interpolates and adds transitional years where there are gaps in the timeseries data.

```{r anim2}
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() + # Convert x to log scale
  transition_time(year)
anim2
```

The much smoother movement in Option 2 will be much more noticeable if you add a title to the chart, that will page through the years corresponding to each frame.


Q5 Can you add a title to one or both of the animations above that will change 
 in sync with the animation? [hint: search labeling for transition_states() and transition_time() functions respectively]
```{r, Q5}
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) + # I choose to title animation two
  geom_point() +
  scale_x_log10() + # Convert x to log scale
  transition_time(year) + 
  labs(title = "Year: {frame_time}") # The title (year) will change in sync with the animation
anim2

```
 

Q6 Can you make the axes' labels and units more readable? Consider expanding the abreviated lables as well as the scientific notation in the legend and x axis to whole numbers.[hint:search disabling scientific notation]
```{r, Q6}
options(scipen = 999) # I found help here: https://stackoverflow.com/questions/5352099/how-to-disable-scientific-notation
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) + # I choose to title animation two
  geom_point() +
  scale_x_log10() + # Convert x to log scale
  transition_time(year) + 
  labs(title = "Year: {frame_time}") # The title (year) will change in sync with the animation
anim2

```


Q7 Come up with a question you want to answer using the gapminder data and write it down. Then, create a data visualisation that answers the question and explain how your visualization answers the question. (Example: you wish to see what was mean life expectancy across the continents in the year you were born versus your parents' birth years). [hint: if you wish to have more data than is in the filtered gapminder, you can load either the `gapminder_unfiltered` dataset and download more at https://www.gapminder.org/data/ ]

I want to answer the following question: How has Denmark's life expectancy developed over time? 

```{r}
gapminder_DK <- gapminder %>%
  select(country, year, lifeExp) %>% 
  filter(country == "Denmark") %>% 
  arrange(desc(lifeExp))
gapminder_DK
```

```{r}
ggplot(subset(gapminder, country == "Denmark"), aes(year, lifeExp)) +
         geom_point()
```

# Q7: The development of life expectancy in Denmark has increased over time from being 70 years in 1952 to being 78 in 2007.



