---
title: 'W7 Homework: Share of female managers'
author: "Sofia Sif Overby Fjorback"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal
I want to be able to analyse OECD.stat's data on share of female managers from 2011-2018 (https://stats.oecd.org/index.aspx?queryid=96330)

# Solution
First, install a handful of classic R packages and load their libraries:

- `rvest` for web-scraping
- `dplyr` for data-wrangling
- `tidyr` for data transformation
- `stringr` for string manipulation
- `janitor` for clean headers that your OCD will love you for
- 'tidyverse' for cleaning data
- 'ggplot2' for vizualisations


```{r libraries}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
library(tidyverse)
library(ggplot2)
```


## Scrape the data

Next, learn how scrape the content of the website and extract the HTML table:
```{r}
url <- "https://stats.oecd.org/index.aspx?queryid=96330"
# scrape the website
url_html <- read_html(url)
```


The following option let's us extract the whole HTML table through the <table> tag.

```{r, option3}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list

```

If you run a `head()` on the resulting `whole_table`, you will see that it is a list with unnamed elements marked by numbers in double brackets, such as [[1]]. 

```{r htmltodf}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table) 
```

The line above takes the whole table from the website with some rows and coloumns. It does not look nice and it shows thing that I am not interested in.

# Tidying the data

In my case I only want to look at the countries' share of female managers through the time period (2011 to 2018). Therefore I filter the data as shown in the following lines

```{r create final table}
femalemanagers <- new_table[8:52,2:18] #I choose the rows and columns I want to look at 
femalemanagers <- femalemanagers[,c(1,3:10)] #I choose the columns I want to both eliminate and look at
colnames(femalemanagers) <- c("country", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018") #I assign new column names to years
femalemanagers
```

# Visualizations 

I want to visualize the development of share of female managers in specific countries

```{r chart of Denmark}
femalemanagers <- as.data.frame(femalemanagers)

femalemanagers %>%
  filter(country == "Denmark")
class(femalemanagers)

fmanagers <- femalemanagers
```


```{r, chart of share of female managers}
fmanagers <- as_tibble(femalemanagers)
fmanagers
```
I want to look at Denmark

```{r, chart Denmark}
fmanagers %>% 
  filter(country == "Denmark")

```

I try to visualize this development

```{r, visualization of Denmark's development}
fmanagers <- femalemanagers %>% 
  filter(country == "Denmark")

ggplot(fmanagers, aes(x = , y =))
geom_point()
```
For some reason it does not work, because the data is not structured in the right way.

Instead I choose to look for the countries that have the highest share of female managers:

```{r, the countries with highest shares in 2011}
fmanagers %>% 
  filter(year == 2011)
  arrange(desc(fmanagers)) %>% 
  head(10)
```

I thought it could be interesting to compare the lists of highest shares in 2011 and 2018 to see if something has changed... To conclude I finished scraping a website and ended up with a  table, but I did not end up visualizing anything, because the table should be structured in another way. This is what I need to practice for next time. 

Question: Could the difficulties with the visualizations have something to do with the fact that R does not understand that the table shows percentages?? 


